---
title: Funding
description: "Competitive research funding supporting the MINDSET programme, including fellowships and collaborative awards from UKRI, ADR UK, NIHR, and other funders."
output:
  html_document:
    keep_md: yes
---

My research is supported through competitive fellowships, investigator-led awards, collaborative programmes, and research-development funding. Funding information is maintained through my [ORCID record](https://orcid.org/0000-0003-2881-4906), with my role and contribution shown where relevant.

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(orcidtr)
library(tidyverse)
library(httr2)
library(lubridate)
library(stringr)
library(purrr)
library(htmltools)

orcid_id <- "0000-0003-2881-4906"

# -----------------------------------------------------------------------------
# Local annotations
# -----------------------------------------------------------------------------

funding_overrides <- read_csv(
  "funding_overrides.csv",
  show_col_types = FALSE,
  col_types = cols(
    grant_number = col_character(),
    match_title = col_character(),
    role = col_character(),
    description_override = col_character(),
    end_date_override = col_character()
  )
) |>
  mutate(
    across(everything(), ~ na_if(str_squish(.x), "")),
    match_title_key = match_title |>
      str_to_lower() |>
      str_replace_all("[[:punct:]]", " ") |>
      str_squish()
  )

# -----------------------------------------------------------------------------
# Retrieve full funding records from ORCID
# -----------------------------------------------------------------------------

fetch_funding_detail <- function(orcid_id, put_code) {

  url <- paste0(
    "https://pub.orcid.org/v3.0/",
    orcid_id,
    "/funding/",
    put_code
  )

  resp <- request(url) |>
    req_headers(Accept = "application/json") |>
    req_error(is_error = \(r) FALSE) |>
    req_perform()

  if (resp_status(resp) != 200) {
    return(list(
      description = NA_character_,
      amount = NA_character_,
      currency = NA_character_,
      url = NA_character_,
      grant_number = NA_character_
    ))
  }

  b <- resp_body_json(resp)

  external_ids <- b$`external-ids`$`external-id`

  grant_number <- if (!is.null(external_ids) && length(external_ids) > 0) {
    ids <- map_chr(
      external_ids,
      \(x) x$`external-id-value` %||% NA_character_
    )
    ids <- ids[!is.na(ids) & str_trim(ids) != ""]
    if (length(ids) > 0) ids[[1]] else NA_character_
  } else {
    NA_character_
  }

  list(
    description = b$`short-description` %||% NA_character_,
    amount = b$amount$value %||% NA_character_,
    currency = b$amount$`currency-code` %||% NA_character_,
    url = b$url$value %||% NA_character_,
    grant_number = grant_number
  )
}

funding_raw <- as_tibble(orcid_funding(orcid_id))

details <- map(
  funding_raw$put_code,
  \(pc) fetch_funding_detail(orcid_id, pc)
)

funding <- funding_raw |>
  mutate(
    description = map_chr(details, "description"),
    amount = map_chr(details, "amount"),
    currency = map_chr(details, "currency"),
    url = map_chr(details, "url"),
    grant_number = map_chr(details, "grant_number")
  )

# -----------------------------------------------------------------------------
# Clean ORCID metadata
# -----------------------------------------------------------------------------

funding <- funding |>
  mutate(
    title_key = title |>
      str_to_lower() |>
      str_replace_all("[[:punct:]]", " ") |>
      str_squish(),

    grant_number = na_if(str_squish(grant_number), ""),

    start_year = suppressWarnings(as.integer(substr(start_date, 1, 4))),
    start_month = suppressWarnings(as.integer(substr(start_date, 6, 7))),
    end_year = suppressWarnings(as.integer(substr(end_date, 1, 4))),
    end_month = suppressWarnings(as.integer(substr(end_date, 6, 7))),

    orcid_role = case_when(
      str_detect(coalesce(description, ""), "^\\s*\\[[^]]+\\]") ~
        str_match(description, "^\\s*\\[([^]]+)\\]")[, 2],
      TRUE ~ NA_character_
    ),

    description_clean = description |>
      str_remove("^\\s*\\[[^]]+\\]\\s*") |>
      str_squish()
  )

# -----------------------------------------------------------------------------
# Apply overrides
#
# Grant number is the preferred key. Title matching is only a fallback.
# -----------------------------------------------------------------------------

overrides_by_grant <- funding_overrides |>
  filter(!is.na(grant_number)) |>
  distinct(grant_number, .keep_all = TRUE) |>
  select(
    grant_number,
    role_grant = role,
    description_grant = description_override,
    end_date_grant = end_date_override
  )

overrides_by_title <- funding_overrides |>
  filter(!is.na(match_title_key)) |>
  distinct(match_title_key, .keep_all = TRUE) |>
  select(
    match_title_key,
    role_title = role,
    description_title = description_override,
    end_date_title = end_date_override
  )

funding <- funding |>
  left_join(overrides_by_grant, by = "grant_number") |>
  left_join(
    overrides_by_title,
    by = c("title_key" = "match_title_key")
  ) |>
  mutate(
    role = coalesce(role_grant, role_title, orcid_role),
    description_display = coalesce(
      description_grant,
      description_title,
      description_clean
    ),
    end_date_display = coalesce(
      end_date_grant,
      end_date_title,
      end_date
    ),
    end_year = suppressWarnings(as.integer(substr(end_date_display, 1, 4))),
    end_month = suppressWarnings(as.integer(substr(end_date_display, 6, 7)))
  )

# -----------------------------------------------------------------------------
# Classify by role
# -----------------------------------------------------------------------------

funding <- funding |>
  mutate(
    funding_group = case_when(
      str_detect(
        coalesce(role, ""),
        regex(
          "(Principal Investigator|\\bPI\\b|Fellow|Fellowship Lead|Programme Lead)",
          ignore_case = TRUE
        )
      ) ~ "lead",

      str_detect(
        coalesce(role, ""),
        regex(
          "(Co-Investigator|Co-I|Work Package Lead|WP Lead|Workstream Lead|Co-Lead)",
          ignore_case = TRUE
        )
      ) ~ "collaborative",

      TRUE ~ "other"
    )
  )

# -----------------------------------------------------------------------------
# Format metadata
# -----------------------------------------------------------------------------

funding <- funding |>
  mutate(
    currency_symbol = recode(
      coalesce(currency, ""),
      GBP = "£",
      USD = "$",
      EUR = "€",
      .default = coalesce(currency, "")
    ),

    amount_numeric = suppressWarnings(as.numeric(amount)),

    amount_fmt = case_when(
      !is.na(amount_numeric) & !is.na(currency) & currency != "" ~
        paste0(
          currency_symbol,
          formatC(
            amount_numeric,
            format = "f",
            digits = 0,
            big.mark = ","
          )
        ),
      TRUE ~ NA_character_
    ),

    start_label = case_when(
      !is.na(start_year) & !is.na(start_month) ~
        paste(month(start_month, label = TRUE, abbr = TRUE), start_year),
      !is.na(start_year) ~ as.character(start_year),
      TRUE ~ ""
    ),

    end_label = case_when(
      !is.na(end_year) & !is.na(end_month) ~
        paste(month(end_month, label = TRUE, abbr = TRUE), end_year),
      !is.na(end_year) ~ as.character(end_year),
      TRUE ~ "Present"
    ),

    date_range = case_when(
      start_label != "" & end_label != "" ~ paste0(start_label, " to ", end_label),
      start_label != "" ~ start_label,
      TRUE ~ ""
    )
  ) |>
  arrange(desc(start_year), desc(start_month))

# -----------------------------------------------------------------------------
# Alert during knitting if anything has no role annotation
# -----------------------------------------------------------------------------

unclassified <- funding |>
  filter(is.na(role) | str_trim(role) == "")

if (nrow(unclassified) > 0) {
  message(
    "\nFunding records without a classified role:\n- ",
    paste(unclassified$title, collapse = "\n- "),
    "\n\nAdd a row to funding_overrides.csv if you want these classified.\n"
  )
}
```


```{r include=FALSE}
render_funding <- function(df) {

  if (nrow(df) == 0) {
    return(invisible(NULL))
  }

  cards <- pmap_chr(
    df,
    function(
      title,
      organization,
      description_display,
      amount_fmt,
      url,
      role,
      date_range,
      grant_number,
      ...
    ) {

      safe_title <- htmlEscape(title)

      title_html <- if (!is.na(url) && str_trim(url) != "") {
        paste0(
          '<div class="funding-title"><a href="',
          htmlEscape(url),
          '">',
          safe_title,
          "</a></div>"
        )
      } else {
        paste0(
          '<div class="funding-title">',
          safe_title,
          "</div>"
        )
      }

      role_html <- if (!is.na(role) && str_trim(role) != "") {
        paste0(
          '<div class="funding-role">',
          htmlEscape(role),
          "</div>"
        )
      } else {
        ""
      }

      meta_parts <- c(
        if (!is.na(organization) && organization != "")
          htmlEscape(organization),
        if (!is.na(amount_fmt) && amount_fmt != "")
          htmlEscape(amount_fmt),
        if (!is.na(date_range) && date_range != "")
          htmlEscape(date_range)
      )

      meta_parts <- meta_parts[!is.na(meta_parts) & meta_parts != ""]

      meta_html <- if (length(meta_parts) > 0) {
        paste0(
          '<div class="funding-meta">',
          paste(meta_parts, collapse = " &middot; "),
          "</div>"
        )
      } else {
        ""
      }

      description_html <- if (
        !is.na(description_display) &&
        str_trim(description_display) != ""
      ) {
        paste0(
          '<p class="funding-description">',
          htmlEscape(description_display),
          "</p>"
        )
      } else {
        ""
      }

      reference_html <- if (
        !is.na(grant_number) &&
        str_trim(grant_number) != ""
      ) {
        paste0(
          '<div class="funding-reference">',
          "Award ",
          htmlEscape(grant_number),
          "</div>"
        )
      } else {
        ""
      }

      paste0(
        '<div class="funding-card">',
        title_html,
        role_html,
        meta_html,
        description_html,
        reference_html,
        "</div>"
      )
    }
  )

  cat(paste(cards, collapse = "\n"))
}
```

## Fellowships and lead investigator awards

<div class="funding-section-intro">
Awards for which I hold primary responsibility for the intellectual direction and delivery of the research.
</div>

```{r results='asis', echo=FALSE}
funding |>
  filter(funding_group == "lead") |>
  render_funding()
```

## Collaborative research awards

<div class="funding-section-intro">
Major collaborative programmes in which I hold a defined investigator or work-package leadership role.
</div>

```{r results='asis', echo=FALSE}
funding |>
  filter(funding_group == "collaborative") |>
  render_funding()
```

## Research development and enabling awards

<div class="funding-section-intro">
Additional funding supporting research infrastructure, public involvement, methodological development, training, and research culture.
</div>

```{r results='asis', echo=FALSE}
funding |>
  filter(funding_group == "other") |>
  render_funding()
```
