---
title: Publications
description: "Selected and complete publications from the MINDSET research programme, spanning psychiatric epidemiology, linked data, mental health data science, neurodivergence, addiction, and health inequalities."
output:
  html_document:
    keep_md: yes
---

My publications span psychiatric epidemiology, linked administrative and clinical data, mental-health inequalities, neurodivergence, addiction, and population health. The programme-defining outputs highlighted below are complemented by a dynamically generated view of recent work and my complete publication record from [ORCID](https://orcid.org/0000-0003-2881-4906).

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(orcidtr)
library(rcrossref)
library(tidyverse)
library(httr2)
library(stringr)
library(purrr)
library(htmltools)

orcid_id <- "0000-0003-2881-4906"

selected <- read_csv(
  "selected_publications.csv",
  show_col_types = FALSE,
  col_types = cols(
    order = col_double(),
    doi = col_character(),
    theme = col_character(),
    significance = col_character(),
    contribution = col_character()
  )
) |>
  mutate(
    doi = str_to_lower(str_squish(doi))
  ) |>
  arrange(order)

w <- orcid_works(orcid_id)

df <- as_tibble(w) |>
  mutate(
    doi = str_to_lower(doi),
    year = as.integer(substr(publication_date, 1, 4)),
    title_key = title |>
      str_to_lower() |>
      str_replace_all("[[:punct:]]", " ") |>
      str_squish(),
    is_preprint = type %in% c("preprint", "posted-content"),
    has_doi = !is.na(doi)
  ) |>
  arrange(title_key, is_preprint, desc(has_doi), desc(year))

with_doi <- df |>
  filter(has_doi)

no_doi <- df |>
  filter(!has_doi)

cr_cites <- with_doi |>
  mutate(
    citation = map_chr(
      doi,
      \(d) {
        tryCatch(
          cr_cn(dois = d, format = "text"),
          error = \(e) NA_character_
        )
      }
    )
  )

fallback <- bind_rows(
  no_doi,
  cr_cites |>
    filter(is.na(citation)) |>
    select(-citation)
) |>
  mutate(
    citation = str_trim(
      paste0(
        title,
        if_else(
          is.na(journal) | journal == "",
          "",
          paste0(". *", journal, "*")
        ),
        if_else(
          is.na(year),
          "",
          paste0(" (", year, ").")
        )
      )
    )
  )

publications <- bind_rows(
  cr_cites |>
    filter(!is.na(citation)),
  fallback
) |>
  mutate(
    citation = if_else(
      str_ends(citation, "\\."),
      citation,
      paste0(citation, ".")
    ),
    citation = str_replace_all(
      citation,
      "Yang,\\s*J\\.\\s*C\\.",
      "**Yang, J. C.**"
    ),
    citation = str_replace_all(
      citation,
      "(?<!C\\.)\\bYang,\\s*J\\.(?!\\s*C)",
      "**Yang, J.**"
    )
  ) |>
  arrange(desc(year), title)

excluded_preprint_dois <- c(
  "10.1101/2024.06.25.24309483",
  "10.1101/2023.03.29.23287850"
)

peer_reviewed <- publications |>
  filter(type == "journal-article") |>
  distinct(title_key, .keep_all = TRUE)

published_dois <- str_to_lower(peer_reviewed$doi)

book_chapters_raw <- publications |>
  filter(type == "book-chapter") |>
  filter(!title_key %in% peer_reviewed$title_key) |>
  distinct(title_key, .keep_all = TRUE)

doi_chapter_titles <- book_chapters_raw |>
  filter(!is.na(doi)) |>
  pull(title_key)

book_chapters <- book_chapters_raw |>
  mutate(
    redundant_no_doi = map_lgl(
      seq_len(n()),
      \(i) {
        if (!is.na(doi[[i]]) || length(doi_chapter_titles) == 0) {
          return(FALSE)
        }

        x <- title_key[[i]]

        any(
          map_lgl(
            doi_chapter_titles,
            \(y) str_detect(x, fixed(y)) || str_detect(y, fixed(x))
          )
        )
      }
    )
  ) |>
  filter(!redundant_no_doi) |>
  select(-redundant_no_doi) |>
  mutate(
    citation = citation |>
      str_replace_all("\\*{3,}", "**") |>
      str_squish()
  )

get_published_doi <- function(doi) {
  tryCatch({
    resp <- request(
      paste0(
        "https://api.crossref.org/works/",
        URLencode(doi, reserved = TRUE)
      )
    ) |>
      req_headers(Accept = "application/json") |>
      req_error(is_error = \(r) FALSE) |>
      req_perform()

    if (resp_status(resp) != 200) {
      return(NA_character_)
    }

    body <- resp_body_json(resp)
    relations <- body$message$relation
    published <- relations$`is-preprint-of` %||%
      relations$`is-version-of`

    if (is.null(published) || length(published) == 0) {
      return(NA_character_)
    }

    published[[1]]$id %||% NA_character_

  }, error = \(e) NA_character_)
}

preprints_raw <- publications |>
  filter(type %in% c("preprint", "posted-content")) |>
  filter(!title_key %in% peer_reviewed$title_key) |>
  distinct(title_key, .keep_all = TRUE)

preprints <- preprints_raw |>
  mutate(
    published_doi = map_chr(
      doi,
      \(d) {
        if (is.na(d)) return(NA_character_)
        get_published_doi(d)
      }
    ),
    has_published_version = (
      !is.na(published_doi) &
        str_to_lower(published_doi) %in% published_dois
    ) |
      str_to_lower(doi) %in%
        str_to_lower(excluded_preprint_dois)
  ) |>
  filter(!has_published_version)

selected_display <- selected |>
  left_join(
    peer_reviewed |>
      select(
        doi,
        title,
        year,
        journal,
        citation
      ),
    by = "doi"
  )

recent_display <- peer_reviewed |>
  filter(!doi %in% selected$doi) |>
  arrange(desc(year), desc(publication_date)) |>
  slice_head(n = 5)
```


## Programme-defining outputs

These papers illustrate the development of the main strands of my research programme and areas in which I have held substantial intellectual or analytical responsibility.

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
selected_display |>
  filter(!is.na(title)) |>
  pmap_chr(
    function(
      order,
      doi,
      theme,
      significance,
      contribution,
      title,
      year,
      journal,
      citation,
      ...
    ) {

      title_html <- paste0(
        '<div class="selected-title">',
        '<a href="https://doi.org/',
        htmlEscape(doi),
        '">',
        htmlEscape(title),
        "</a></div>"
      )

      meta_parts <- c(
        if (!is.na(journal) && journal != "")
          htmlEscape(journal),
        if (!is.na(year))
          as.character(year)
      )

      meta_html <- paste0(
        '<div class="selected-meta">',
        paste(meta_parts, collapse = " &middot; "),
        "</div>"
      )

      theme_html <- if (
        !is.na(theme) &&
        str_trim(theme) != ""
      ) {
        paste0(
          '<div class="selected-theme">',
          htmlEscape(theme),
          "</div>"
        )
      } else {
        ""
      }

      significance_html <- if (
        !is.na(significance) &&
        str_trim(significance) != ""
      ) {
        paste0(
          "<p>",
          htmlEscape(significance),
          "</p>"
        )
      } else {
        ""
      }

      contribution_html <- if (
        !is.na(contribution) &&
        str_trim(contribution) != ""
      ) {
        paste0(
          '<p class="selected-contribution"><b>My contribution:</b> ',
          htmlEscape(contribution),
          "</p>"
        )
      } else {
        ""
      }

      paste0(
        '<div class="selected-publication">',
        theme_html,
        title_html,
        meta_html,
        significance_html,
        contribution_html,
        "</div>"
      )
    }
  ) |>
  paste(collapse = "\n") |>
  cat()
```

## Recent peer-reviewed work

The most recent peer-reviewed publications not already highlighted above are shown automatically from ORCID.

```{r results='asis', echo=FALSE}
recent_display |>
  mutate(
    doi_link = if_else(
      !is.na(doi),
      paste0("https://doi.org/", doi),
      NA_character_
    )
  ) |>
  pmap_chr(
    function(title, journal, year, doi_link, ...) {
      title_html <- if (!is.na(doi_link)) {
        paste0(
          '<a href="',
          htmlEscape(doi_link),
          '"><b>',
          htmlEscape(title),
          "</b></a>"
        )
      } else {
        paste0("<b>", htmlEscape(title), "</b>")
      }

      meta <- paste(
        c(
          if (!is.na(journal) && journal != "") htmlEscape(journal),
          if (!is.na(year)) as.character(year)
        ),
        collapse = " · "
      )

      paste0(
        '<div class="recent-publication"><p>',
        title_html,
        if (meta != "") paste0("<br><span class=\"selected-meta\">", meta, "</span>") else "",
        "</p></div>"
      )
    }
  ) |>
  paste(collapse = "\n") |>
  cat()
```

## Complete publication list

<div class="publication-list">

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat("### Peer-reviewed publications\n\n")
cat(
  paste0(peer_reviewed$citation),
  sep = "\r  \n"
)

if (nrow(book_chapters) > 0) {
  cat("\n\n### Book chapters\n\n")
  cat(
    paste0(book_chapters$citation),
    sep = "\r  \n"
  )
}

if (nrow(preprints) > 0) {
  cat("\n\n### Preprints\n\n")
  cat(
    paste0(preprints$citation),
    sep = "\r  \n"
  )
}
```

</div>
