---
title: Publications
description: "Selected and complete publications underpinning MINDSET's work in psychiatric epidemiology, multimodal inference, mental health data science, neurodivergence, addiction, and health inequalities."
output:
  html_document:
    keep_md: yes
---

My publications reflect the programme's current empirical foundations in psychiatric epidemiology, linked administrative and clinical data, mental-health inequalities, neurodivergence, addiction, spatial analysis, and population health. These foundations support a wider multimodal agenda in which substantive questions determine when complementary evidence should be integrated across levels, settings, and timescales. The programme-defining outputs highlighted below are complemented by a dynamically generated view of recent work and my complete publication record from [ORCID](https://orcid.org/0000-0003-2881-4906).

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(orcidtr)
library(rcrossref)
library(tidyverse)
library(httr2)
library(stringr)
library(purrr)
library(htmltools)

source("../../R/publications.R")

orcid_id <- "0000-0003-2881-4906"

selected <- read_csv(
  "../../data/site/selected_publications.csv",
  show_col_types = FALSE,
  col_types = cols(
    order = col_double(),
    doi = col_character(),
    theme = col_character(),
    significance = col_character(),
    contribution = col_character()
  )
) |>
  mutate(
    doi = normalize_doi(doi)
  ) |>
  arrange(order)

w <- orcid_works(orcid_id)

df <- as_tibble(w) |>
  mutate(
    doi = normalize_doi(doi),
    year = as.integer(substr(publication_date, 1, 4)),
    title_key = title |>
      str_to_lower() |>
      str_replace_all("[[:punct:]]", " ") |>
      str_squish(),
    is_preprint = type %in% c("preprint", "posted-content"),
    has_doi = !is.na(doi)
  ) |>
  arrange(title_key, is_preprint, desc(has_doi), desc(year))

with_doi <- df |>
  filter(has_doi)

no_doi <- df |>
  filter(!has_doi)

crossref_cache_path <- "../../data/generated/crossref-bibtex.json"

cr_cites <- with_doi |>
  left_join(
    refresh_crossref_citations(with_doi$doi, crossref_cache_path),
    by = "doi"
  )

fallback <- bind_rows(
  no_doi,
  cr_cites |>
    filter(is.na(citation)) |>
    select(-citation)
) |>
  mutate(
    citation = str_trim(
      paste0(
        title,
        if_else(
          is.na(journal) | journal == "",
          "",
          paste0(". *", journal, "*")
        ),
        if_else(
          is.na(year),
          "",
          paste0(" (", year, ").")
        )
      )
    )
  )

publications <- bind_rows(
  cr_cites |>
    filter(!is.na(citation)),
  fallback
) |>
  mutate(
    citation = if_else(
      str_ends(citation, "\\."),
      citation,
      paste0(citation, ".")
    ),
    citation = str_replace_all(
      citation,
      "Yang,\\s*J\\.\\s*C\\.",
      "**Yang, J. C.**"
    ),
    citation = str_replace_all(
      citation,
      "(?<!C\\.)\\bYang,\\s*J\\.(?!\\s*C)",
      "**Yang, J.**"
    )
  ) |>
  arrange(desc(year), title)

excluded_preprint_dois <- c(
  "10.1101/2024.06.25.24309483",
  "10.1101/2023.03.29.23287850"
)

peer_reviewed <- publications |>
  filter(type == "journal-article") |>
  distinct(title_key, .keep_all = TRUE)

published_dois <- str_to_lower(peer_reviewed$doi)

book_chapters_raw <- publications |>
  filter(type == "book-chapter") |>
  filter(!title_key %in% peer_reviewed$title_key) |>
  distinct(title_key, .keep_all = TRUE)

doi_chapter_titles <- book_chapters_raw |>
  filter(!is.na(doi)) |>
  pull(title_key)

book_chapters <- book_chapters_raw |>
  mutate(
    redundant_no_doi = map_lgl(
      seq_len(n()),
      \(i) {
        if (!is.na(doi[[i]]) || length(doi_chapter_titles) == 0) {
          return(FALSE)
        }

        x <- title_key[[i]]

        any(
          map_lgl(
            doi_chapter_titles,
            \(y) str_detect(x, fixed(y)) || str_detect(y, fixed(x))
          )
        )
      }
    )
  ) |>
  filter(!redundant_no_doi) |>
  select(-redundant_no_doi) |>
  mutate(
    citation = citation |>
      str_replace_all("\\*{3,}", "**") |>
      str_squish()
  )

get_published_doi <- function(doi) {
  tryCatch({
    resp <- request(
      paste0(
        "https://api.crossref.org/works/",
        URLencode(doi, reserved = TRUE)
      )
    ) |>
      req_headers(Accept = "application/json") |>
      req_error(is_error = \(r) FALSE) |>
      req_perform()

    if (resp_status(resp) != 200) {
      return(NA_character_)
    }

    body <- resp_body_json(resp)
    relations <- body$message$relation
    published <- relations$`is-preprint-of` %||%
      relations$`is-version-of`

    if (is.null(published) || length(published) == 0) {
      return(NA_character_)
    }

    published[[1]]$id %||% NA_character_

  }, error = \(e) NA_character_)
}

preprints_raw <- publications |>
  filter(type %in% c("preprint", "posted-content")) |>
  filter(!title_key %in% peer_reviewed$title_key) |>
  distinct(title_key, .keep_all = TRUE)

preprints <- preprints_raw |>
  mutate(
    published_doi = map_chr(
      doi,
      \(d) {
        if (is.na(d)) return(NA_character_)
        get_published_doi(d)
      }
    ),
    has_published_version = (
      !is.na(published_doi) &
        str_to_lower(published_doi) %in% published_dois
    ) |
      str_to_lower(doi) %in%
        str_to_lower(excluded_preprint_dois)
  ) |>
  filter(!has_published_version)

publication_sections <- bind_rows(
  peer_reviewed |>
    mutate(publication_section = "peer-reviewed"),
  book_chapters |>
    mutate(publication_section = "book-chapter"),
  preprints |>
    mutate(publication_section = "preprint")
) |>
  distinct(title_key, publication_section, .keep_all = TRUE) |>
  add_publication_ids()

openalex_enrichment <- refresh_openalex_enrichment(
  publication_sections$doi,
  "../../data/generated/openalex-publications.json"
)

selected_strand_overrides <- selected |>
  transmute(
    doi,
    selected_strand = classify_selected_strand(theme)
  )

publication_sections <- publication_sections |>
  left_join(openalex_enrichment, by = "doi") |>
  left_join(selected_strand_overrides, by = "doi") |>
  mutate(
    journal = coalesce(
      na_if(journal, ""),
      openalex_source
    ),
    automatic_strand = classify_research_strand(
      title,
      openalex_topics
    ),
    research_strand = coalesce(
      selected_strand,
      automatic_strand
    ),
    research_strand_label = unname(
      research_strand_labels[research_strand]
    ),
    open_access_url = if_else(
      is_open_access %in% TRUE &
        !is.na(open_access_url) &
        open_access_url != "",
      open_access_url,
      NA_character_
    ),
    publication_search = str_to_lower(
      paste(
        coalesce(title, ""),
        coalesce(journal, ""),
        coalesce(as.character(year), ""),
        coalesce(research_strand_label, ""),
        coalesce(openalex_topics, "")
      )
    )
  ) |>
  select(-selected_strand, -automatic_strand)

bibtex_records <- generate_bibtex_records(
  publication_sections,
  crossref_cache_path
)

writeLines(
  paste(bibtex_records$bibtex, collapse = "\n\n"),
  "justin-yang-publications.bib",
  useBytes = TRUE
)

publication_sections <- publication_sections |>
  left_join(bibtex_records, by = "publication_id")

peer_reviewed <- publication_sections |>
  filter(publication_section == "peer-reviewed") |>
  arrange(desc(year), title)

book_chapters <- publication_sections |>
  filter(publication_section == "book-chapter") |>
  arrange(desc(year), title)

preprints <- publication_sections |>
  filter(publication_section == "preprint") |>
  arrange(desc(year), title)

bibtex_json <- bibtex_records |>
  select(
    publication_id,
    bibtex,
    bibtex_filename
  ) |>
  jsonlite::toJSON(
    dataframe = "rows",
    auto_unbox = TRUE,
    na = "null"
  ) |>
  str_replace_all("</", "<\\/")

selected_display <- selected |>
  left_join(
    peer_reviewed |>
      select(
        doi,
        title,
        year,
        journal,
        citation
      ),
    by = "doi"
  )

```


## Programme-defining outputs

These papers illustrate the development of the main strands of my research programme and areas in which I have held substantial intellectual or analytical responsibility.

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
selected_display |>
  filter(!is.na(title)) |>
  pmap_chr(
    function(
      order,
      doi,
      theme,
      significance,
      contribution,
      title,
      year,
      journal,
      citation,
      ...
    ) {

      title_html <- paste0(
        '<div class="selected-title">',
        '<a href="https://doi.org/',
        htmlEscape(doi),
        '">',
        htmlEscape(title),
        "</a></div>"
      )

      meta_parts <- c(
        if (!is.na(journal) && journal != "")
          htmlEscape(journal),
        if (!is.na(year))
          as.character(year)
      )

      meta_html <- paste0(
        '<div class="selected-meta">',
        paste(meta_parts, collapse = " &middot; "),
        "</div>"
      )

      theme_html <- if (
        !is.na(theme) &&
        str_trim(theme) != ""
      ) {
        paste0(
          '<div class="selected-theme">',
          htmlEscape(theme),
          "</div>"
        )
      } else {
        ""
      }

      significance_html <- if (
        !is.na(significance) &&
        str_trim(significance) != ""
      ) {
        paste0(
          "<p>",
          htmlEscape(significance),
          "</p>"
        )
      } else {
        ""
      }

      contribution_html <- if (
        !is.na(contribution) &&
        str_trim(contribution) != ""
      ) {
        paste0(
          '<p class="selected-contribution"><b>My contribution:</b> ',
          htmlEscape(contribution),
          "</p>"
        )
      } else {
        ""
      }

      paste0(
        '<div class="selected-publication">',
        theme_html,
        title_html,
        meta_html,
        significance_html,
        contribution_html,
        "</div>"
      )
    }
  ) |>
  paste(collapse = "\n") |>
  cat()
```

## Explore my research

The timeline and filters organise my publications into broad research strands using OpenAlex topics, with curated classifications retained for the programme-defining outputs above. Open-access links are included where OpenAlex identifies an accessible copy.

<div id="publication-explorer" class="publication-explorer">

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
strand_buttons <- c(
  '<button type="button" class="publication-filter" data-strand-filter="all" aria-pressed="true">All strands</button>',
  imap_chr(
    research_strand_labels,
    function(label, strand) {
      paste0(
        '<button type="button" class="publication-filter publication-strand-',
        htmlEscape(strand, attribute = TRUE),
        '" data-strand-filter="',
        htmlEscape(strand, attribute = TRUE),
        '" aria-pressed="false">',
        '<span class="publication-strand-dot" aria-hidden="true"></span>',
        htmlEscape(label),
        '</button>'
      )
    }
  )
)

cat(
  '<div class="publication-filter-group" aria-label="Filter publications by research strand">',
  paste(strand_buttons, collapse = "\n"),
  '</div>',
  sep = "\n"
)

timeline_data <- peer_reviewed |>
  filter(
    !is.na(year),
    !is.na(research_strand)
  )

if (nrow(timeline_data) > 0) {
  plot_left <- 165
  plot_right <- 875
  plot_top <- 48
  lane_gap <- 42
  timeline_height <- plot_top +
    (length(research_strand_labels) - 1) * lane_gap +
    32

  first_year <- min(timeline_data$year)
  last_year <- max(timeline_data$year)
  year_span <- max(last_year - first_year, 1)
  x_for_year <- function(x) {
    plot_left +
      ((x - first_year) / year_span) *
      (plot_right - plot_left)
  }

  year_breaks <- pretty(
    c(first_year, last_year),
    n = 8
  ) |>
    as.integer() |>
    keep(\(x) x >= first_year && x <= last_year) |>
    unique()

  if (!first_year %in% year_breaks) {
    year_breaks <- c(first_year, year_breaks)
  }
  if (!last_year %in% year_breaks) {
    year_breaks <- c(year_breaks, last_year)
  }

  lane_positions <- set_names(
    plot_top +
      (seq_along(research_strand_labels) - 1) *
      lane_gap,
    names(research_strand_labels)
  )

  timeline_data <- timeline_data |>
    group_by(year, research_strand) |>
    arrange(title, .by_group = TRUE) |>
    mutate(
      mark_number = row_number(),
      marks_in_cell = n(),
      mark_offset = (
        mark_number -
          (marks_in_cell + 1) / 2
      ) * pmin(10, 42 / pmax(marks_in_cell, 1))
    ) |>
    ungroup() |>
    mutate(
      mark_x = x_for_year(year) + mark_offset,
      mark_y = unname(lane_positions[research_strand])
    )

  vertical_grid <- map_chr(
    year_breaks,
    function(year_value) {
      x <- x_for_year(year_value)
      paste0(
        '<line class="publication-timeline-grid" x1="',
        round(x, 2),
        '" y1="28" x2="',
        round(x, 2),
        '" y2="',
        timeline_height - 12,
        '"></line>',
        '<text class="publication-timeline-year" x="',
        round(x, 2),
        '" y="18" text-anchor="middle">',
        year_value,
        '</text>'
      )
    }
  )

  horizontal_grid <- imap_chr(
    research_strand_labels,
    function(label, strand) {
      y <- lane_positions[[strand]]
      paste0(
        '<line class="publication-timeline-grid" x1="',
        plot_left,
        '" y1="',
        y,
        '" x2="',
        plot_right,
        '" y2="',
        y,
        '"></line>',
        '<text class="publication-timeline-lane" x="0" y="',
        y + 4,
        '">',
        htmlEscape(label),
        '</text>'
      )
    }
  )

  timeline_marks <- timeline_data |>
    pmap_chr(
      function(
        publication_id,
        title,
        journal,
        year,
        research_strand,
        research_strand_label,
        mark_x,
        mark_y,
        ...
      ) {
        label <- c(
          title,
          if (!is.na(journal) && journal != "") journal,
          year
        ) |>
          paste(collapse = " · ")
        paste0(
          '<circle class="publication-timeline-mark publication-strand-',
          htmlEscape(research_strand, attribute = TRUE),
          '" cx="',
          round(mark_x, 2),
          '" cy="',
          round(mark_y, 2),
          '" r="6" role="button" tabindex="0" data-publication-id="',
          htmlEscape(publication_id, attribute = TRUE),
          '" data-label="',
          htmlEscape(label, attribute = TRUE),
          '" aria-label="',
          htmlEscape(label, attribute = TRUE),
          '"></circle>'
        )
      }
    )

  cat(
    '<div class="publication-timeline-scroll">',
    paste0(
      '<svg class="publication-timeline" viewBox="0 0 900 ',
      timeline_height,
      '" role="img" aria-labelledby="publication-timeline-title publication-timeline-description">'
    ),
    '<title id="publication-timeline-title">Publication timeline by research strand</title>',
    '<desc id="publication-timeline-description">Each point represents a peer-reviewed publication. Select a point to identify the paper.</desc>',
    paste(vertical_grid, collapse = "\n"),
    paste(horizontal_grid, collapse = "\n"),
    paste(timeline_marks, collapse = "\n"),
    '</svg>',
    '</div>',
    '<p id="publication-selected-paper" class="publication-selected-paper" aria-live="polite">Select a point to identify the publication.</p>',
    sep = "\n"
  )
}
```

## Complete publication list

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
publication_years <- publication_sections$year |>
  stats::na.omit() |>
  unique() |>
  sort(decreasing = TRUE)

year_options <- c(
  '<option value="all">All years</option>',
  map_chr(
    publication_years,
    \(year_value) paste0(
      '<option value="',
      year_value,
      '">',
      year_value,
      '</option>'
    )
  )
)

cat(
  '<div class="publication-toolbar">',
  '<label class="publication-control" for="publication-search">Search publications',
  '<input id="publication-search" type="search" placeholder="Title, journal or topic">',
  '</label>',
  '<label class="publication-control" for="publication-year">Publication year',
  '<select id="publication-year">',
  paste(year_options, collapse = "\n"),
  '</select>',
  '</label>',
  '<a class="publication-action publication-download-all" href="justin-yang-publications.bib" download>Export all (.bib)</a>',
  '</div>',
  paste0(
    '<p id="publication-result-count" class="publication-result-count" aria-live="polite">',
    nrow(publication_sections),
    ' publications</p>'
  ),
  sep = "\n"
)

render_publication_section <- function(data, heading) {
  if (nrow(data) == 0) {
    return(invisible(NULL))
  }

  entries <- data |>
    pmap_chr(
      function(
        publication_id,
        title,
        journal,
        year,
        doi,
        citation,
        research_strand,
        research_strand_label,
        publication_search,
        is_open_access,
        open_access_url,
        open_access_status,
        ...
      ) {
        doi_action <- if (!is.na(doi) && doi != "") {
          paste0(
            '<a class="publication-action" href="https://doi.org/',
            htmlEscape(doi, attribute = TRUE),
            '">DOI</a>'
          )
        } else {
          ""
        }

        open_access_duplicates_doi <-
          !is.na(doi) &&
          !is.na(open_access_url) &&
          normalize_doi(open_access_url) == normalize_doi(doi)

        open_access_action <- if (
          !is.na(open_access_url) &&
          open_access_url != "" &&
          !open_access_duplicates_doi
        ) {
          paste0(
            '<a class="publication-action" href="',
            htmlEscape(open_access_url, attribute = TRUE),
            '">Open-access copy</a>'
          )
        } else {
          ""
        }

        open_access_badge <- if (is_open_access %in% TRUE) {
          '<span class="publication-open-access">Open access</span>'
        } else {
          ""
        }

        paste0(
          '<article id="',
          htmlEscape(publication_id, attribute = TRUE),
          '" class="publication-entry publication-strand-',
          htmlEscape(research_strand, attribute = TRUE),
          '" data-strand="',
          htmlEscape(research_strand, attribute = TRUE),
          '" data-year="',
          if_else(is.na(year), "", as.character(year)),
          '" data-search="',
          htmlEscape(publication_search, attribute = TRUE),
          '">',
          '<div class="publication-citation">',
          citation_to_html(citation),
          '</div>',
          '<div class="publication-entry-meta">',
          '<span class="publication-strand-label"><span class="publication-strand-dot" aria-hidden="true"></span>',
          htmlEscape(research_strand_label),
          '</span>',
          open_access_badge,
          '</div>',
          '<div class="publication-entry-actions">',
          doi_action,
          open_access_action,
          '<button type="button" class="publication-action" data-bibtex-id="',
          htmlEscape(publication_id, attribute = TRUE),
          '">BibTeX</button>',
          '</div>',
          '</article>'
        )
      }
    )

  cat(
    '<section class="publication-section" data-publication-section>',
    paste0('<h3>', htmlEscape(heading), '</h3>'),
    paste(entries, collapse = "\n"),
    '</section>',
    sep = "\n"
  )
}

render_publication_section(
  peer_reviewed,
  "Peer-reviewed publications"
)
render_publication_section(
  book_chapters,
  "Book chapters"
)
render_publication_section(
  preprints,
  "Preprints"
)

cat(
  '<script id="publication-bibtex-data" type="application/json">',
  bibtex_json,
  '</script>',
  sep = "\n"
)
```

</div>
