---
title: Talks & Presentations
description: "Invited talks, seminars, conference presentations, workshops, and methods training by Dr Justin C Yang."
output:
  html_document:
    keep_md: yes
---

I regularly present work in psychiatric epidemiology, linked data, mental health data science, neurodivergence, addiction, and research methods to academic, policy, and interdisciplinary audiences.

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(glue)
library(htmltools)
library(lubridate)

talks <- read_csv(
  "talks.csv",
  show_col_types = FALSE,
  col_types = cols(
    Host = col_character(),
    Title = col_character(),
    Event = col_character(),
    Date = col_date(),
    Type = col_character(),
    Invited = col_logical(),
    Materials = col_character()
  )
) |>
  mutate(
    Type = na_if(str_squish(Type), ""),
    Materials = na_if(str_squish(Materials), "")
  )

today <- Sys.Date()

upcoming <- talks |>
  filter(Date >= today) |>
  arrange(Date)

recent <- talks |>
  filter(
    Date < today,
    Date >= today - years(3)
  ) |>
  arrange(desc(Date))

older <- talks |>
  filter(Date < today - years(3)) |>
  arrange(desc(Date))

render_talks <- function(df) {

  if (nrow(df) == 0) {
    return(invisible(NULL))
  }

  cards <- pmap_chr(
    df,
    function(
      Host,
      Title,
      Event,
      Date,
      Type,
      Invited,
      Materials,
      ...
    ) {

      labels <- c(
        if (!is.na(Invited) && Invited) "Invited",
        if (!is.na(Type) && Type != "") Type
      )

      label_html <- if (length(labels) > 0) {
        paste0(
          '<div class="talk-label">',
          htmlEscape(paste(labels, collapse = " · ")),
          "</div>"
        )
      } else {
        ""
      }

      title_html <- paste0(
        '<div class="talk-title">',
        htmlEscape(Title),
        "</div>"
      )

      event_parts <- c(
        if (!is.na(Event) && Event != "") htmlEscape(Event),
        if (!is.na(Host) && Host != "") htmlEscape(Host)
      )

      event_html <- paste0(
        '<div class="talk-event">',
        paste(event_parts, collapse = " · "),
        "</div>"
      )

      date_html <- paste0(
        '<div class="talk-date">',
        format(Date, "%d %B %Y"),
        "</div>"
      )

      materials_html <- if (
        !is.na(Materials) &&
        Materials != ""
      ) {
        paste0(
          '<div class="talk-materials"><a href="',
          htmlEscape(Materials),
          '">View presentation materials</a></div>'
        )
      } else {
        ""
      }

      paste0(
        '<div class="talk-card">',
        label_html,
        title_html,
        event_html,
        date_html,
        materials_html,
        "</div>"
      )
    }
  )

  cat(paste(cards, collapse = "\n"))
}
```

<style>
.talk-card {
  padding: 1.05rem 0 1.2rem 0;
  border-bottom: 1px solid #e6e6e6;
}

.talk-card:last-child {
  border-bottom: none;
}

.talk-label {
  font-size: 0.8rem;
  font-weight: 600;
  text-transform: uppercase;
  letter-spacing: 0.035em;
  opacity: 0.7;
  margin-bottom: 0.2rem;
}

.talk-title {
  font-size: 1.04rem;
  font-weight: 600;
  line-height: 1.4;
  margin-bottom: 0.25rem;
}

.talk-event {
  font-size: 0.93rem;
  line-height: 1.45;
  margin-bottom: 0.15rem;
}

.talk-date {
  font-size: 0.87rem;
  opacity: 0.72;
}

.talk-materials {
  font-size: 0.9rem;
  margin-top: 0.35rem;
}
</style>

```{r results='asis', echo=FALSE}
if (nrow(upcoming) > 0) {
  cat("## Upcoming\n\n")
  render_talks(upcoming)
}
```

## Recent talks and presentations

```{r results='asis', echo=FALSE}
render_talks(recent)
```

```{r results='asis', echo=FALSE}
if (nrow(older) > 0) {
  cat("\n\n<details>\n")
  cat("<summary><b>Earlier talks and presentations</b></summary>\n\n")
  render_talks(older)
  cat("\n\n</details>\n")
}
```
