---
title: Talks & Presentations
description: "Talks, seminars, conference presentations, workshops, and methods training by Dr Justin C Yang."
output:
  html_document:
    keep_md: yes
---

I present research from across MINDSET on mental-health trajectories and inequalities, linked and heterogeneous evidence, neurodivergence, addiction, and the causal, spatial, and data-science methods that support this work. This page lists selected talks, presentations, workshops, and methods training for academic, policy, clinical, service, and interdisciplinary audiences.

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(htmltools)
library(lubridate)

talks <- read_csv(
  "../../data/site/talks.csv",
  show_col_types = FALSE,
  col_types = cols(
    Host = col_character(),
    Title = col_character(),
    Event = col_character(),
    Date = col_date(),
    Type = col_character(),
    Invited = col_logical(),
    Materials = col_character()
  )
) |>
  mutate(
    Type = na_if(str_squish(Type), ""),
    Materials = na_if(str_squish(Materials), "")
  )

today <- Sys.Date()

upcoming <- talks |>
  filter(Date >= today) |>
  arrange(Date)

recent <- talks |>
  filter(
    Date < today,
    Date >= today - years(3)
  ) |>
  arrange(desc(Date))

older <- talks |>
  filter(Date < today - years(3)) |>
  arrange(desc(Date))

render_talks <- function(df) {

  if (nrow(df) == 0) {
    return(invisible(NULL))
  }

  cards <- pmap_chr(
    df,
    function(
      Host,
      Title,
      Event,
      Date,
      Type,
      Invited,
      Materials,
      ...
    ) {

      labels <- c(
        if (!is.na(Invited) && Invited) "Invited",
        if (!is.na(Type) && Type != "") Type
      )

      label_html <- if (length(labels) > 0) {
        paste0(
          '<div class="talk-label">',
          htmlEscape(paste(labels, collapse = " · ")),
          "</div>"
        )
      } else {
        ""
      }

      title_html <- paste0(
        '<div class="talk-title">',
        htmlEscape(Title),
        "</div>"
      )

      event_parts <- c(
        if (!is.na(Event) && Event != "") htmlEscape(Event),
        if (!is.na(Host) && Host != "") htmlEscape(Host)
      )

      event_html <- paste0(
        '<div class="talk-event">',
        paste(event_parts, collapse = " · "),
        "</div>"
      )

      date_html <- paste0(
        '<div class="talk-date">',
        format(Date, "%d %B %Y"),
        "</div>"
      )

      materials_html <- if (
        !is.na(Materials) &&
        Materials != ""
      ) {
        link_text <- if (!is.na(Type) && Type == "Course") {
          "View course"
        } else {
          "View presentation materials"
        }

        paste0(
          '<div class="talk-materials"><a href="',
          htmlEscape(Materials),
          '">',
          link_text,
          "</a></div>"
        )
      } else {
        ""
      }

      paste0(
        '<div class="talk-card">',
        label_html,
        title_html,
        event_html,
        date_html,
        materials_html,
        "</div>"
      )
    }
  )

  cat(paste(cards, collapse = "\n"))
}
```

```{r results='asis', echo=FALSE}
if (nrow(upcoming) > 0) {
  cat("## Upcoming\n\n")
  render_talks(upcoming)
}
```

## Recent talks and presentations

```{r results='asis', echo=FALSE}
render_talks(recent)
```

```{r results='asis', echo=FALSE}
if (nrow(older) > 0) {
  cat("\n\n<details>\n")
  cat("<summary><b>Earlier talks and presentations</b></summary>\n\n")
  render_talks(older)
  cat("\n\n</details>\n")
}
```
