---
title: Teaching & Supervision
description: "Teaching, supervision, trainee development, and quantitative methods training by Dr Justin C Yang at UCL."
output:
  html_document:
    keep_md: yes
---

My teaching and supervision are organised around a simple aim: **helping trainees become thoughtful, independent researchers who can reason clearly with complex health and population data**. I teach quantitative methods, supervise research across mental health and population health, and contribute to open and reproducible research training.

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(stringr)

students <- read_csv(
  "../../data/site/students.csv",
  show_col_types = FALSE
)

# Keep a complete portable supervision ledger alongside the curated page.
write_csv(students, "supervision-record.csv", na = "")

current_phd_students <- students |>
  filter(
    Degree == "PhD",
    str_detect(`Year(s)`, "Present"),
    str_detect(str_to_lower(Role), "supervisor")
  )

current_phd_committees <- students |>
  filter(
    Degree == "PhD",
    str_detect(`Year(s)`, "Present"),
    str_detect(str_to_lower(Role), "committee")
  )

current_taught <- students |>
  filter(
    Degree != "PhD",
    str_detect(`Year(s)`, "Present")
  )

primary_projects <- students |>
  filter(
    str_to_lower(Role) == "primary supervisor",
    Degree != "PhD"
  )

published_projects <- students |>
  filter(
    Degree != "PhD",
    str_detect(Topic, "\\]\\(https?://")
  )

recent_projects <- students |>
  filter(
    Institution == "UCL",
    Degree != "PhD",
    !str_detect(`Year(s)`, "Present")
  ) |>
  arrange(desc(`Year(s)`), Student) |>
  slice_head(n = 8)
```

## Current trainees

### PhD students

I currently contribute as secondary supervisor to PhD work spanning neurodevelopment, educational experiences, and inequalities in psychiatric disorders.

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
current_phd_students |>
  arrange(Student) |>
  select(Student, Role, Topic) |>
  knitr::kable(
    format = "html",
    col.names = c("PhD student", "Role", "Research"),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

### Thesis committees

I also contribute to doctoral research through thesis-committee roles where my methodological or substantive expertise is relevant.

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
current_phd_committees |>
  arrange(Student) |>
  select(Student, Topic) |>
  knitr::kable(
    format = "html",
    col.names = c("PhD student", "Research"),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

### MSc supervision

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
if (nrow(current_taught) > 0) {
  current_taught |>
    arrange(Student) |>
    select(Student, Degree, Role, Topic) |>
    knitr::kable(
      format = "html",
      col.names = c("Student", "Degree", "Role", "Project"),
      table.attr = 'class="site-table"'
    ) |>
    cat()
}
```

</div>

## Teaching & training

At UCL, I teach epidemiology, statistics, and research methods across postgraduate mental-health programmes. Current and recent teaching includes:

<div class="teaching-module-grid">
<div class="teaching-module"><svg class="teaching-module-glyph" viewBox="0 0 64 40" aria-hidden="true" focusable="false"><circle cx="32" cy="6" r="4"/><path d="M32 10v8M10 18h44M10 18v10m22-10v10m22-10v10"/><circle cx="10" cy="32" r="4"/><circle cx="32" cy="32" r="4"/><circle cx="54" cy="32" r="4"/></svg><a href="https://www.ucl.ac.uk/module-catalogue/modules/core-principles-of-mental-health-research-PSBS0002"><b>Core Principles of Mental Health Research</b></a><br>Research design, critical appraisal, and the foundations of mental-health research.</div>
<div class="teaching-module"><svg class="teaching-module-glyph" viewBox="0 0 64 40" aria-hidden="true" focusable="false"><path d="M7 34V6M7 34h51M12 29l10-6 9 2 10-10 12-5"/><circle cx="15" cy="16" r="2.5"/><circle cx="26" cy="30" r="2.5"/><circle cx="36" cy="18" r="2.5"/><circle cx="49" cy="22" r="2.5"/></svg><a href="https://www.ucl.ac.uk/module-catalogue/modules/practical-statistics-for-mental-health-research-PSBS0008"><b>Practical Statistics for Mental Health Research</b></a><br>Applied quantitative reasoning and statistical analysis for mental-health research.</div>
<div class="teaching-module"><svg class="teaching-module-glyph" viewBox="0 0 64 40" aria-hidden="true" focusable="false"><circle cx="8" cy="11" r="4"/><circle cx="8" cy="29" r="4"/><path d="M12 11h18l9 9h17M12 29h18l9-9M48 14l8 6-8 6"/></svg><a href="https://www.ucl.ac.uk/module-catalogue/modules/epidemiological-research-methods-in-mental-health-PSBS0013"><b>Epidemiological Research Methods in Mental Health</b></a><br>Epidemiological designs, measures, inference, and interpretation in psychiatric research.</div>
</div>

I have also contributed to curriculum development around **open science** and **public involvement and engagement with research**. As a certified instructor with [The Carpentries](https://carpentries.org/), I teach practical data and computational skills for open and reproducible research. Selected reusable teaching materials are available through [Open Research & Resources](../resources/).

## Supervision in practice

```{r results='asis', echo=FALSE}
cat(
  '<div class="supervision-summary">',
  '<div class="supervision-stat"><b>', nrow(primary_projects), '</b><span>primary-supervised postgraduate or undergraduate projects</span></div>',
  '<div class="supervision-stat"><b>', nrow(current_phd_students), '</b><span>current PhD students</span></div>',
  '<div class="supervision-stat"><b>', nrow(current_phd_committees), '</b><span>current thesis-committee roles</span></div>',
  '<div class="supervision-stat"><b>', nrow(published_projects), '</b><span>student projects linked to published outputs</span></div>',
  '</div>'
)
```

The supervision portfolio spans mental health, addiction, neurodivergence, climate and health, health inequalities, public health, and health policy. Several projects have developed into peer-reviewed outputs.

### Student work linked to publications

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
published_projects |>
  arrange(desc(`Year(s)`), Student) |>
  select(Student, Degree, Role, Topic) |>
  knitr::kable(
    format = "html",
    col.names = c("Student", "Degree", "Role", "Published work"),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

### Selected recent supervision

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
recent_projects |>
  select(Student, Degree, Role, Topic) |>
  knitr::kable(
    format = "html",
    col.names = c("Student", "Degree", "Role", "Project"),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

## Prospective trainees

I am interested in supervising PhD, MSc, and undergraduate projects that fit the wider MINDSET agenda, particularly in psychiatric epidemiology, linked administrative or clinical data, causal inference, mental-health data science, neurodivergence, addiction, and inequalities. Prospective trainees are welcome to [get in touch](../contact/) with a brief outline of their interests and the type of project or supervision they are seeking.
