---
title: Teaching & Supervision
description: "Researcher development, supervision, teaching, and quantitative methods training by Dr Justin C Yang at UCL."
output:
  html_document:
    keep_md: yes
---

My teaching and supervision focus on **developing researchers who can work critically, confidently, and responsibly with complex health and population data**. This includes doctoral and master's supervision, quantitative methods teaching, open and reproducible research training, and wider mentoring of early-career researchers.

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(stringr)

students <- read_csv(
  "../../data/site/students.csv",
  show_col_types = FALSE
)

current_phd <- students |>
  filter(
    Degree == "PhD",
    str_detect(`Year(s)`, "Present")
  )

primary_projects <- students |>
  filter(
    str_to_lower(Role) == "primary supervisor",
    Degree != "PhD"
  )

published_projects <- students |>
  filter(
    Degree != "PhD",
    str_detect(Topic, "\\]\\(https?://")
  )
```

## Researcher development & supervision

I currently contribute to the supervision and development of doctoral researchers across neurodevelopment, health inequalities, population mental health, learning disability, and the effects of extreme heat on severe mental illness.

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
current_phd |>
  arrange(Student) |>
  select(
    Student,
    Role,
    Topic
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Researcher",
      "Role",
      "Research"
    ),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

Alongside doctoral supervision, I have substantial experience as a primary and secondary supervisor for master's and undergraduate research. The supervision record below is generated from a maintained project list; several student projects have subsequently contributed to peer-reviewed outputs.

```{r results='asis', echo=FALSE}
cat(
  '<div class="supervision-summary">',
  '<b>', nrow(primary_projects), '</b> primary-supervised postgraduate or undergraduate projects',
  ' &middot; ',
  '<b>', nrow(current_phd), '</b> current doctoral supervision or thesis-committee roles',
  ' &middot; ',
  '<b>', nrow(published_projects), '</b> listed student projects linked to published outputs',
  '</div>'
)
```

I also mentor and support early-career researchers through collaborative projects, methods training, research-culture initiatives, and informal research development. I am a certified instructor with [The Carpentries](https://carpentries.org/), with a focus on practical data and computational skills for open and reproducible research.

## Postgraduate teaching

At UCL, I teach epidemiology, statistics, and research methods across postgraduate mental-health programmes. Teaching has included:

- [Core Principles of Mental Health Research](https://www.ucl.ac.uk/module-catalogue/modules/core-principles-of-mental-health-research-PSBS0002)
- [Practical Statistics for Mental Health Research](https://www.ucl.ac.uk/module-catalogue/modules/practical-statistics-for-mental-health-research-PSBS0008)
- [Epidemiological Research Methods in Mental Health](https://www.ucl.ac.uk/module-catalogue/modules/epidemiological-research-methods-in-mental-health-PSBS0013)

I have also contributed to curriculum development around **open science** and **public involvement and engagement with research**.

My approach to quantitative teaching emphasises reasoning about evidence: what a study can and cannot support, when stronger or causal conclusions are justified, how analytical decisions influence results, and how findings can be communicated clearly and responsibly. I use applied examples from epidemiology and mental-health research and place particular emphasis on making quantitative methods accessible to students with different levels of prior experience.

## MSc and undergraduate research supervision

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
students |>
  filter(Degree != "PhD") |>
  arrange(
    desc(`Year(s)`),
    Student
  ) |>
  select(
    Student,
    Degree,
    `Year(s)`,
    Role,
    Topic
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Student",
      "Degree",
      "Year(s)",
      "Role",
      "Project"
    ),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

## Previous doctoral committee role

<div class="site-table-wrap">

```{r results='asis', echo=FALSE}
doctoral_previous <- students |>
  filter(
    Degree == "PhD",
    !str_detect(`Year(s)`, "Present")
  ) |>
  arrange(
    desc(`Year(s)`),
    Student
  ) |>
  select(
    Student,
    `Year(s)`,
    Role,
    Topic
  )

if (nrow(doctoral_previous) > 0) {
  doctoral_previous |>
    knitr::kable(
      format = "html",
      col.names = c(
        "Researcher",
        "Year(s)",
        "Role",
        "Research"
      ),
      table.attr = 'class="site-table"'
    ) |>
    cat()
}
```

</div>

Prospective PhD, MSc, and undergraduate students interested in psychiatric epidemiology, causal inference, mental-health data science, or related areas are welcome to [get in touch](../contact/) to discuss potential supervision or project ideas.
