---
title: Teaching & Supervision
description: "Teaching, doctoral supervision, researcher development, and quantitative methods training by Dr Justin C Yang at UCL."
output:
  html_document:
    keep_md: yes
---

My teaching and supervision focus on **epidemiology, quantitative methods, open and reproducible research, and the development of early-career researchers**. I teach primarily at postgraduate level and supervise students working across psychiatric epidemiology, mental health data science, causal inference, public health, and related areas.

## Doctoral supervision

I currently contribute to the supervision of doctoral researchers working on neurodevelopment, psychiatric inequalities, and population mental health.

<div class="site-table-wrap">

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(stringr)

students <- read_csv(
  "students.csv",
  show_col_types = FALSE
)

doctoral_current <- students |>
  filter(
    Degree == "PhD",
    str_detect(`Year(s)`, "Present")
  ) |>
  arrange(Student) |>
  select(
    Student,
    Role,
    Topic
  )

if (nrow(doctoral_current) > 0) {
  doctoral_current |>
    knitr::kable(
      format = "html",
      col.names = c(
        "Researcher",
        "Supervisory role",
        "Research"
      ),
      table.attr = 'class="site-table"'
    ) |>
    cat()
}
```

</div>

I also serve on doctoral thesis committees, contributing methodological and substantive expertise to projects beyond those I formally co-supervise.

## Researcher development

Developing researchers is an important part of my academic work. My contributions include mentoring and supporting early-career researchers, developing training in quantitative and reproducible research methods, and creating opportunities for researchers to build confidence, independence, and collaborative networks.

I am a certified instructor with [The Carpentries](https://carpentries.org/), an international organisation focused on building practical data and computational skills for open and reproducible research.

My teaching and researcher-development work also includes contributions to open science, public involvement and engagement, and research culture within UCL and wider collaborative programmes.

## Postgraduate teaching

At UCL, I teach epidemiology, statistics, and research methods across postgraduate mental-health programmes. Teaching has included:

- [Core Principles of Mental Health Research](https://www.ucl.ac.uk/module-catalogue/modules/core-principles-of-mental-health-research-PSBS0002)
- [Practical Statistics for Mental Health Research](https://www.ucl.ac.uk/module-catalogue/modules/practical-statistics-for-mental-health-research-PSBS0008)
- [Epidemiological Research Methods in Mental Health](https://www.ucl.ac.uk/module-catalogue/modules/epidemiological-research-methods-in-mental-health-PSBS0013)

I have also contributed to curriculum development around **open science** and **public involvement and engagement with research**.

My approach to quantitative teaching emphasises careful reasoning about evidence: what a study can and cannot support, when stronger or causal conclusions are justified, how analytical decisions influence results, and how findings can be communicated clearly and responsibly. I use applied examples from epidemiology and mental-health research and place particular emphasis on making quantitative methods accessible to students with different levels of prior experience.

## MSc and undergraduate supervision

I have supervised a wide range of student research projects across psychiatric epidemiology, public health, mental health, health inequalities, substance use, policy, and quantitative methods. Several supervised projects have contributed to peer-reviewed publications.

<div class="site-table-wrap">

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
non_doctoral <- students |>
  filter(Degree != "PhD") |>
  arrange(
    desc(`Year(s)`),
    Student
  ) |>
  select(
    Student,
    Degree,
    `Year(s)`,
    Role,
    Topic
  )

non_doctoral |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Student",
      "Degree",
      "Year(s)",
      "Role",
      "Project"
    ),
    table.attr = 'class="site-table"'
  ) |>
  cat()
```

</div>

## Previous doctoral committee roles

<div class="site-table-wrap">

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
doctoral_previous <- students |>
  filter(
    Degree == "PhD",
    !str_detect(`Year(s)`, "Present")
  ) |>
  arrange(
    desc(`Year(s)`),
    Student
  ) |>
  select(
    Student,
    `Year(s)`,
    Role,
    Topic
  )

if (nrow(doctoral_previous) > 0) {
  doctoral_previous |>
    knitr::kable(
      format = "html",
      col.names = c(
        "Researcher",
        "Year(s)",
        "Role",
        "Research"
      ),
      table.attr = 'class="site-table"'
    ) |>
    cat()
}
```

</div>

Prospective PhD, MSc, and undergraduate students interested in psychiatric epidemiology, causal inference, mental health data science, or related areas are welcome to [get in touch](../contact/) to discuss potential supervision or project ideas.
